---
title: "ERDDAP data into R Notebook"
output: html_notebook
---

Download a ncdf file from erddap.  Generate the url on the erddap and paste it in here between the single quotes
The code below will download a few years of chl data from the OC-CCI for a region around San Francisco and Monterey Bay.  

```{r}
test <- download.file('https://coastwatch.pfeg.noaa.gov/erddap/griddap/pmlEsaCCI31OceanColorMonthly.nc?chlor_a[(2008-01-01):1:(2018-06-01T00:00:00Z)][(38.5):1:(36.5)][(-123.5):1:(-121.5)]',destfile='test.nc',mode='wb')
```

Now read the ncdf file that was downloaded by the request above  

```{r}
datafileID <- nc_open('test.nc')

chlor_a   <- ncvar_get(datafileID, "chlor_a")
longitude <- ncvar_get(datafileID, "longitude")
latitude  <- ncvar_get(datafileID, "latitude")
datatime  <- ncvar_get(datafileID, "time")

nc_close(datafileID)
```

Convert the time variable into a seconds since 1970 to a variable recognized as a date by R 

```{r}
datatime <- as.Date(as.POSIXlt(datatime,origin='1970-01-01',tz= "GMT"))
```

Average the 2D data into a timeseries of the average chlorophyll value. 
The c(3) indicates the time dimension will be the one dimension returned    

```{r}
AvgChl <- apply(chlor_a, c(3),function(x) mean(x,na.rm=TRUE))
```

Plot the timeseries 

```{r}
plot(datatime, AvgChl,type='b', pch=20)
```
Average the data into a map the average chlorophyll value. 
The c(1,2) indicates the x and y dimensions will be returned   

```{r}
AvgChlMap <- apply(chlor_a,c(1,2),function(x) mean(x,na.rm=TRUE))
```

Make a map of the average chl field using base graphics

```{r}
image(longitude,latitude,AvgChlMap)
```
So let's look at the order of values in longitude and latitude.  

```{r}
longitude
```
```{r}
latitude
```
So latitude is the culprit, having descending values. So we will need to reverse this array and rotate the satellite image as well

```{r}

fliplatitude <- sort(latitude)
FlipAvgChlMap <- apply(t(AvgChlMap),2,rev)
image(longitude,fliplatitude,log(FlipAvgChlMap),col=rev(rainbow(12)))
```

THe map is rather distorrted, and it would be nice to get the coastline and color legend on the map.  A lot of these features are easier to do using ggplot, so lets make a map in ggplot. 
First need to restructure the data to plot using ggplot, create a function to do this.  
This function will vectorize the 2D array of satellite data and create vectors of latitude and longitude that are the same length 

```{r}
MapFrame <- function(longitude,latitude,data){
dims<-dim(data)
data<-array(data,dims[1]*dims[2])
MapFrame<-expand.grid(x=longitude,y=latitude)
MapFrame$data<-data
return(MapFrame)
}
```

Now restructure the data by passing it through the function defined above. Pass in the log of the Chlorophyll data  
```{r}
ChlMap <- MapFrame(longitude,latitude,log(AvgChlMap))
```

Call the ggplot2 library

```{r}
library(ggplot2)
```

```{r}
coast <- map_data("worldHires", ylim = range(latitude), xlim = range(longitude))
ggplot(data = ChlMap, aes(x = x, y = y, fill = data)) + 
  geom_raster(interpolate = FALSE,na.rm=T) +
  coord_fixed(1.3,ylim = range(latitude), xlim = range(longitude)) +
  geom_polygon(data = coast, aes(x=long, y = lat, group = group), fill = "grey80") +
  scale_fill_gradientn(colours = rev(rainbow(12))) + 
  ylab("Latitude") + xlab("Longitude") 
```
test <- download.file('http://localhost:8080/erddap/tabledap/BGC_Argo.htmlTable?time%2Clatitude%2Clongitude&time%3E=2012-05-01T00%3A00%3A00Z&time%3C=2012-06-30T00%3A00%3A00Z&latitude%3E=20&latitude%3C=24&longitude%3E=193&longitude%3C=198',destfile='test.nc',mode='wb')
